---
title: "Activity #10: Quarto"
format:
  html:
    math: true
---

## Q1: Plot \\(h(x) \\; for \\; x \\in \[0,1\]\\).

Considering the function \$\$h(x) = \\big\[\\cos(4x)+\\tan(0.5x)\\big\]\^4\$\$

```{r}
#| echo: false
#| include: true
#| message: false
#| warning: false
library(ggplot2)
myFunction <- function(x) {
  y <- (cos(4 * x) + tan(0.5 * x))^4
  return(y)
}
ggplot() +
  stat_function(
    fun = myFunction,
    xlim = c(0, 1),
    n = 1000
  ) +
  theme_bw() +
  labs(
    x = "x",
    y = "h(x)"
  )

```

# Q2: Diamonds Plot

```{r}
#| echo: true
#| include: true
#| message: false
#| warning: false
data(diamonds)
ggplot(diamonds, aes(x = carat, y = price, color = clarity)) +
  geom_point(alpha = 0.5, size = 1.5) +
  labs(
    title = "Relationship Between Diamond Carat and Price",
    subtitle = "Colored by Clarity Level",
    x = "Carat (Weight)",
    y = "Price (USD)",
    color = "Clarity"
  ) +
  scale_color_brewer(palette = "Dark2") +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5),
    legend.position = "right",
    legend.title = element_text(face = "bold")
  )
```
# Q3: Tortilla Prices in Mexico

I did include a **init** for libraries and data here.

```{r, error = TRUE}
#| echo: false
#| include: false
library(tidyverse)
library(readr)
library(knitr)
library(dplyr)
tortilla_data <- read_csv("/Users/aiden/Downloads/q/tortilla_data.csv")
```

a\. Finding Average Price for 2007.

```{r, error = TRUE}
#| echo: true
#| include: true
#| message: false
#| warning: false
avg_2007 <- tortilla_data %>%
  filter(year == 2007) %>%
  summarise(avg_price = mean(price, na.rm = TRUE))

cat("The average price of tortillas in 2007 was:", round(avg_2007$avg_price, 2), "pesos\n")
```
b\. Finding Average Price for 2024.

```{r, error = TRUE}
#| echo: true
#| include: true
#| message: false
#| warning: false
avg_2024 <- tortilla_data %>%
  filter(year == 2024) %>%
  summarise(avg_price = mean(price, na.rm = TRUE))

cat("The average price of tortillas in 2024 was:", round(avg_2024$avg_price, 2), "pesos\n")

```
c\. Find the percent change in the average price of tortillas across all six states together from the year 2007 to 2024

```{r, error = TRUE}
#| echo: true
#| include: true
percent_change <- ((avg_2024$avg_price - avg_2007$avg_price) / avg_2007$avg_price) * 100
cat("The average percent change of tortillas:", percent_change, "pesos\n")
```

d\. Make a plot that shows the average price of tortillas (across all six states together) each year from the years 2007 to 2024.

```{r, error = TRUE}
#| echo: true
#| include: true
yearly_avg <- tortilla_data %>%
  group_by(year) %>%
  summarise(avg_price = mean(price, na.rm = TRUE))

ggplot(yearly_avg, aes(x = year, y = avg_price)) +
  geom_line(color = "blue", linewidth = 1) + 
  geom_point(color = "red", size = 2) +
  labs(
    title = "Average Tortilla Prices Over Time (2007–2024)",
    x = "Year", y = "Average Price (MXN)"
  ) +
theme_minimal()
```

e\. For the year 2016, provide summary statistics for each of the six states’ price of tortillas. Make a table of the average price of tortillas in each state in 2016. Make side-by-side boxplots for each state in 2016.

```{r, error = TRUE}
#| echo: true
#| include: true
summary_2016 <- tortilla_data %>%
  filter(year == 2016) %>%
  group_by(state) %>%
  summarise(
    avg_price = mean(price, na.rm = TRUE),
    median_price = median(price, na.rm = TRUE),
    min_price = min(price, na.rm = TRUE),
    max_price = max(price, na.rm = TRUE),
    price_sd = sd(price, na.rm = TRUE)
  )

kable(summary_2016, caption = "Summary Statistics for Tortilla Prices by State in 2016")

ggplot(tortilla_data %>%
         filter(year == 2016), aes(x = state, y = price)) +
  geom_boxplot(aes(color = state), fill = "grey", alpha = 0.5) +
  labs(
    title = "Distribution of Tortilla Prices by State in 2016",
    subtitle = "Side-by-side Boxplots for Each State",
    x = "State", y = "Price (MXN)"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(face = "bold", hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5)
  )

```
